---
title: "Working with factors (version 0.1)"
author: "Christopher Brown"
date: "January 29, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# I. Libraries and data acquisition  

```{r libraries, message = F, warning = F}
library(tidyverse) # suggested library for this challenge, but many of these problems can be solved without it
# include other libraries used in your solutions here
```

```{r data_import}
# Load data if desired - one option is https://www.gapminder.org/data/
# Otherwise use mtcars, a built-in dataset
```

# II. Problems  

Using the built-in dataset mtcars (or a dataset of your choosing which has data conducive to working with factors, e.g., character strings or categorical variables):  

1. Check whether a vector or columnn is a factor
2. Create a factor
3. Check the factor levels. What do the levels represent?
4. Order factor levels by another variable
5. Order factor levels by first appearance of each label (label refers to the actual value rather than the level)
6. Move the last two factor levels to the beginning
7. Collapse the `make` values that appear once into an `other` category. Optionally, set your own threshold for `other`.
8. Create a new factor column using the distinct values in `make` and another column with categorical values, such as `cyl`.
9. Sort the dataframe by factor levels
10. Filter the data to remove observations of `fiat`, then drop any unused factor levels. What could happen if this isn't done?
11. Recode or relabel the factor values. For example, you might invert, capitalize, or trim the labels in some way.
12. Using `ggplot`, make two bar charts and compare the results when using `make` as a character vector vs a factor. How does the ordering of bars change with each approach?
13. Using `make` as a charcter vector within a dataframe, coerce `make` to a factor and order it appropriately only within a `ggplot` function call.

Other suggested challenges:
14. Within a faceted ggplot, appropriately order factor levels by a continuous variable.
15. Using a grouped dataframe with factor levels, create a ggplot. What are possible issues that can arise?

# III. Possible solutions  

In this section, try to solve one or more of these problems. There are likely to be multiple solutions. Why might one approach be preferred over another?

```{r solution1}
# You may want to create additional code chunks for various problems, especially those involving plotting.

## Hint for getting started with mtcars: 
# Create a new column 'make' using mtcars
df <- mtcars %>% 
  mutate(make = str_extract(rownames(.), pattern = "^\\w+"))
```

```{r ggplot_example}
# This is just a brief refresher on ggplot syntax. Note that this is a flawed plot. Why?
ggplot(df, aes(x = make, y = mpg)) + 
  geom_col() + # consider using geom_bar(), depending on the intended result
  coord_flip()
```


# IV. Feedback  

1. What additional questions do you have about factors, or factor related topics would you like to see addressed?
2. What functions or resources did you use that were most beneficial in solving these problems?
3. Do you have an interesting example or approach that you'd like to share?
